### 2017-04-20 flsa.tex
first commit

\documentclass{article}
\input{pack.tex}
% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Finite Time Analysis of Linear Stochastic Approximation}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
%  David S.~Hippocampus\thanks{Use footnote for providing further
%    information about author (webpage, alternative
%    address)---\emph{not} for acknowledging funding agencies.} \\
%  Department of Computer Science\\
%  Cranberry-Lemon University\\
%  Pittsburgh, PA 15213 \\
%  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\section{Notation}

\section{Problem Setup}
We consider linear stochastic approximation schemes given as
\begin{align}
\theta_t=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),
\end{align}
where we assume
\begin{assumption}
\item $A_t=A+M_t$ and $b_t=b+N_t$, where $\{M_t \in \R^{\ncn},\,t\geq 0\}$ and $\{N_t\in\R^n,\,t\geq 0\}$ are $i.i.d$ sequences such that $\E[M_t]=0$ and $\E[N_t]=0$.
\item There exists a $\ts\in \R^n$ such that $\ts=A^{-1}b$.
\item All the eigenvalues of $A$ have strictly positive real parts.
\item $\E[M_t^\top M_t\ts]=\sigma_1^2$ and $\E[\norm{N_t}^2]=\sigma_2^2$.
\end{assumption}

\begin{example}[Linear Least Squares]
\end{example}

\begin{example}[TD(0)]
\end{example}

\section{Definitions}
\begin{definition}[Weakly Admissible]
\end{definition}

\begin{definition}[Admissible]
\end{definition}

\begin{definition}[Strongly Admissible]
\end{definition}

\begin{lemma}\label{pdt}
Given a AS matrix $A$, there exists a non-singular matrix $U\in \C^{\ncn}$ such that $A=U\Lambda U^{-1}$ where $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
\end{lemma}
Note that in \cref{pdt} the transformation $U\Lambda U^{-1}$ is not unique.
\begin{example}
\end{example}

\begin{definition}
Define symmetric matrices
$
\SD \eqdef(\T_U(A)^\dag+\T_U(A))-\alpha \T_U(A)^\dag \T_U(A),\,
\SR \eqdef(\T_U(A)^\dag+\T_U(A))-\alpha \E[\T_U(A_1)^\dag \T_U(A_1)],\,
$
and constants
$\rhoD \eqdef \inf_{\norm{x}=1} \ip{x,\SD x},\, \rhoR \eqdef \inf_{\norm{x}=1} \ip{x,\SR x},\,$ and $\bu\eqdef \norm{U}^2\norm{U^{-1}}^2 \big(1+\frac2{\alpha\rhoD}\big)\frac{1}{\alpha \rhoR}$
\end{definition}
\begin{lemma}
Given any AS matrix $A$, there exists $\alpha_D>0,\,\alpha_R>0$ and a non-singluar matrix $U$ such that
\begin{enumerate}
\item $\rhoD>0,\,\forall \alpha \in (0,\alpha_D)$.
\item $\rhoR>0,\,\forall \alpha \in (0,\alpha_R)$.
\end{enumerate}
\end{lemma}
\begin{corollary}
There exists $\beta\eqdef\inf_{U : det(U)\neq 0,\alpha\in (0,\alpha_R)} \bu$.
\end{corollary}
\section{Main Results}
\begin{theorem}
\begin{align}
\E[\norm{\zb_t}^2]
\leq \beta
\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{e_0})}{t+1} \right)\,.
\end{align}
\end{theorem}

\section{Comparison With Prior Works}

\subsection{Stochastic Approximation}

\section{Interesting Special Cases}

\section{Application in Reinforcement Learning}
\subsection{RL Setting}

\subsection{TD(0) Algorithm}

\subsection{GTD}
\textbf{Flaws with GTD-MP:}

\textbf{Why is GTD slow:}

\subsection{iLSTD}

\subsection{Comparison of TD, GTD and iLSTD}

\section{Conclusion}

\appendix

\input{linal}

\end{document}