### 2017-04-22 flsa.tex
initial

\documentclass{article}
\input{pack.tex}
% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Finite Time Analysis of Linear Stochastic Approximation}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
%  David S.~Hippocampus\thanks{Use footnote for providing further
%    information about author (webpage, alternative
%    address)---\emph{not} for acknowledging funding agencies.} \\
%  Department of Computer Science\\
%  Cranberry-Lemon University\\
%  Pittsburgh, PA 15213 \\
%  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\section{Notation}

\section{Problem Setup}
We consider linear stochastic approximation schemes given as
\begin{align}
\theta_t=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),
\end{align}
where we assume
\begin{assumption}\label{ass:lsa}
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item $A_t=A+M_t$ and $b_t=b+N_t$, where $\{M_t \in \R^{\ncn},\,t\geq 0\}$ and $\{N_t\in\R^n,\,t\geq 0\}$ are $i.i.d$ sequences such that $\E[M_t]=0$ and $\E[N_t]=0$.
\item There exists a $\ts\in \R^n$ such that $\ts=A^{-1}b$.
\item All the eigenvalues of $A$ have strictly positive real parts.
\item $\E[M_t^\top M_t\ts]=\sigma_1^2$ and $\E[\norm{N_t}^2]=\sigma_2^2$.
\end{itemize}
\end{assumption}

\begin{example}[Linear Least Squares]
\end{example}

\begin{example}[TD(0)]
\end{example}

\section{Notation}
Let $\R^{\ncn}$ and $\C^{\ncn}$ denote the set of $n\times n$ real and complex matrices respectively. Let $\gln\subset\C^{\ncn}$ denote the set of all invertible $\ncn$ complex matrices.

\section{Definitions}
\begin{definition}[Asymptotically Stable]
An $\ncn$ matrix said to be \emph{asymptotically stable} iff all its eigenvalues have positive real parts.
\end{definition}

\begin{definition}\label{def:dist}[Distributions]
Given a distribution $P$ over $\C^{\ncn}$ and let
\begin{align*}
A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
\end{align*}
and let $\P$ denote the set of probability distributions on $\C^{\ncn}$.
\end{definition}

\begin{definition}[Similarity Transformation]
Given any $U\in \gln$, define $\T_U \colon \C^{\ncn}\rightarrow \C^{\ncn}$ be the linear map that takes $A\in \C^{\ncn}$ to $\T_U(A)=U^{-1}\,A\,U$.
\end{definition}
\begin{definition}[Transformed Measure]
Given a distribution $P$ over $\C^{\ncn}$ define the distribution $P_U$ induced by the linear transformation $\T_U$ as follows:
\begin{align*}
\int_{S} dP_U(M) \eqdef \int_{\T_U^{-1}(S)} dP(M)
\end{align*}
\end{definition}

\begin{comment}
\begin{definition}[Induced Distribution]
Let $P\in \P$ be any distribution over $\R^{\ncn}$ and let $p_c^U$ denote the probability distribution over $\C^{\ncn}$ induced by the linear map $\T_U$ such that the pullback of $p_c^U$ using $\T_U$ is $p$.
\end{definition}
\end{comment}

\begin{definition}[Spectral Norms]
Define $\rhod{P}\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\big(A_P+A_P^\dag)-\alpha A_P^\dag A_P)x}$, $\rhos{P}\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\big(A_P+A_P^\dag-\alpha C_P\big)x}$
\end{definition}

\begin{definition}[Weakly Admissible]
A set of distributions $\P$ over $\C^{\ncn}$ is said to be \emph{weakly admissible} if for any given $P\in \P,\,\exists$ a matrix $U_P \in \gln$, real-valued constants $\alphapd>0$ and $\alphaps>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alphapd)$ and $\rhos{P_U}>0,\,\forall \alpha\in(0,\alphaps)$.
\end{definition}
\begin{definition}
Given a weakly admissible class $\P$ and any $P\in\P$, $U\in \gln$ and $\alpha>0$ such that $\rhos{P_{U}}>0$ define
\begin{align*}
\bu&\eqdef \norm{U}^2\norm{U^{-1}}^2 \left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{1}{\alpha \rhos{P_U}}\\
\beta(P)&\eqdef \inf_{\alpha,U\colon \bu>0} \bu
\end{align*}
\end{definition}

\begin{definition}[Admissible]
A set of distributions $\P$ over $\C^{\ncn}$ is \emph{admissible} if  for any given $P\in \P,\,\exists$ a matrix $U_P\in \gln,\,$ and a real-valued constant $\alpha_0>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alpha_0)$ and $\rhos{P_{U_P}},\,\forall \alpha\in(0,\alpha_0)$.
\end{definition}
\begin{example}
\end{example}
\begin{definition}[Strongly Admissible]
A set of distributions $\P$ over $\C^{\ncn}$ is \emph{strongly admissible} if  $\forall\,P\in \P,\,\exists U_P\in \gln,\, \alpha_0>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alpha_0]$ and $\rhos{P_{U_P}},\,\forall \alpha\in(0,\alpha_0]$, $\inf_{P\in \P} \rho_d(\alpha_0,P_{U_P})>0$ and $\inf_{P\in \P} \rho_s(\alpha_0,P_{U_P})>0$.
\end{definition}
\begin{example}
\end{example}
\begin{comment}
\begin{lemma}\label{pdt}
Given a AS matrix $A$, there exists a non-singular matrix $U\in \C^{\ncn}$ such that $A=U\Lambda U^{-1}$ where $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
\end{lemma}
Note that in \cref{pdt} the transformation $U\Lambda U^{-1}$ is not unique.
\begin{example}
\end{example}
\end{comment}
\begin{comment}
\begin{lemma}
Given any AS matrix $A$, there exists $\alpha_D>0,\,\alpha_R>0$ and a non-singluar matrix $U$ such that
\begin{enumerate}
\item $\rhod{\Lambda_1}>0,\,\forall \alpha \in (0,\alpha_D)$.
\item $\rhos{\Lambda_1}>0,\,\forall \alpha \in (0,\alpha_R)$.
\end{enumerate}
\end{lemma}
\begin{corollary}
There exists $\beta\eqdef\inf_{U : det(U)\neq 0,\alpha\in (0,\alpha_R)} \bu$.
\end{corollary}
\end{comment}
\section{Main Results}
The following theorem states that weak admissibility implies stabilizability.
\begin{theorem}
Let $\P_{as}$ be a set of distributions supported on $\R^{\ncn}$ such that for any given $P\in \P$, all the eigenvalues of $A_{P}$ have positive real parts. $\P_{as}$ is \emph{weakly admissible}.
\end{theorem}
In short, so long as the matrices $A_t$ are distributed such that all the eigenvalues of $\E[A_t]$ have with positive real parts, then there always exists a positive step size $\alpha$ (albeit problem dependent) that results in \eqref{lsa} to converge in the mean-squared sense.

In \Cref{thm:rate} that we state we below, we let $P_A$ denote the underlying distribution that generates $A_t, t\geq 0$, and in order to be consitent with \Cref{ass:lsa} and \Cref{def:dist} it is understood that $A_{P_A}=A$.
\begin{theorem}\label{thm:rate}
\begin{align}
\E[\norm{\zb_t}^2]
\leq \beta(P_A)
\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{e_0})}{t+1} \right)\,.
\end{align}

\end{theorem}

\section{Comparison With Prior Works}

\subsection{Stochastic Approximation}

\section{Interesting Special Cases}

\section{Application in Reinforcement Learning}
\subsection{RL Setting}

\subsection{TD(0) Algorithm}

\subsection{GTD}
\textbf{Flaws with GTD-MP:}

\textbf{Why is GTD slow:}

\subsection{iLSTD}

\subsection{Comparison of TD, GTD and iLSTD}

\section{Conclusion}

\appendix

\input{linal}

\end{document}