### 2017-05-11 instance.tex

\section{Main Results}\label{sec:mainresults}
In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$. Here, we only present the results followed by a discussion and the detailed proofs can be found in \Cref{sec:appendix}.
\begin{lemma}\label{lm:pd}
Let $P$ be positive definite, then there exists an $\alpha_P>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P$ be Hurwitz, then there exists a $U\in \gln$ such that $P_U$ is positive definite.
\end{lemma}
\begin{comment}
\begin{proof}
\begin{align*}
\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x\\
&\geq \lambda_{\min}(A^*_P+A_P)-\alpha \norm{A_P}^2-\sigma^2_P
\end{align*}
The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*_P+A_P)}{\norm{A_P}^2+\sigma^2_P}$
\end{proof}
\end{comment}

\begin{comment}
\begin{theorem}\label{th:pdrate}
Let $P^A$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\ts})}{t+1} \right)\,.
\end{align*}

\end{theorem}
\begin{theorem}\label{thm:simtran}[Change of Basis]
Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\ts$, and suppose if $P_U$ is positive definite, then
\begin{align*}
\EE{\norm{\gh_t-\gamma^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
\end{align*}
where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
\end{theorem}
\end{comment}

\begin{theorem}\label{th:rate}
Let $P$ be Hurwitz, then there exists $U\in\gln$ and $\alpha_{P_U}>0$ such that
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
\end{align*}
\end{theorem}

\begin{theorem}[Lower Bound]
There exists a $P$ Hurwitz, $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
\begin{align*}
\EE{\norm{\eh_t}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \alpha^2 \sigma^2_b \Big)
\end{align*}

\begin{comment}
\begin{align}
\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
\end{align}
\end{comment}
\end{theorem}

\begin{comment}

Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$. Note that in this example $\ts=0$.
By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
\begin{align*}
\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
\end{align*}

\begin{align*}
\norm{\eh_t}^2&=\frac{1}{(t+1)^2}\norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1}\right)e_0}^2
&\leq\frac{1}{(t+1)^2}\alpha^2\norm{A_P^{-1}}^2\norm{e_0}^2
\end{align*}

\begin{align*}
\norm{\eh_t}^2\geq \eh^2_t(1)=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)
\end{align*}
Note that in this example, $\rhos{P^0}=\rhod{P^0}=\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(1-\alpha \lambda_{\min})$. Sandwiching the error between the lower and upper bounds we have
\begin{align}
\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
\end{align}
\end{comment}

\subsection{Result Discussion}
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straighforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$. However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}). This can be remedied by a similiarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$. To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in $\gamma_t$ by pre-multiplying \eqref{constste[} by $U^{-1}$ as shown below:
\begin{align*}
U^{-1}\theta_t&=U^{-1}\theta_{t-1}+\alpha(U^{-1} b_t- U^{-1}A_t U U^{-1}\theta_{t-1}),\\
\gamma_t&=\gamma_{t-1}+\alpha(U^{-1} b_t- J_t\gamma_{t-1}),
\end{align*}
This completes the transformation of a Hurwirz problem $P$ to a positive definite problem $P_U$. One can revert to $\theta_t$ by using $\theta_t=U\gamma_t$, and this explains the reason for the condition number $\cond{U}$ in \Cref{th:rate}. We would like to add that when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}. We would like to mention that such similarity transformations have been in the past for analyzing LSA algorithms in RL \cite{lihong} (see \Cref{sec:related} for more details).
\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} When $P$ is positive definite, we can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{cross} terms of$\EE{\ip{e_i,e_j}},i\neq j$. The growth of the diagonal terms and the cross terms depends on the spectral norm of the random matrix $H_t=I-\alpha A_t$ and the deterministic matrix $H_P=I-\alpha A_P$ respectively. These are given by
\begin{comment}
The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$. The error dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.
\begin{align}\label{eq:errec}
\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1})\nn\\
\theta_t-\ts&=\theta_{t-1}-\ts+\alpha\left(b_t-A_t(\theta_{t-1}-\ts+\ts)\right)\nn\\
e_t&=e_{t-1}+\alpha(b_t-A_t e_t -A_t\ts)\nn\\
e_t&=\underbrace{((I-\alpha A_t))}_{\text{Random-Matrix}} e_{t-1}+\underbrace{\alpha(N_t -(M_t)\ts)}_{\text{Noise}}
\end{align}
From \eqref{errec} it is clear that the error dynamics depends on the $i)$ properties of the random matrix, $ii)$ properties of the noise. In the absence of the noise term in \eqref{eq:errec}, we have $e_t=(I-\alpha A_t) e_{t-1}=\Pi_{s=1}(I-\alpha A_s) e_0$, i.e., the inital error is get multiplied by a product of random matrices. In such a scenario, we can guess that whether or not the algorithm forgets the bias $\norm{\theta_0-\ts}$, depends on whether the matrix product is contracting. One way to characterize this contracting property is to look at the spectral radius of the random matrix, which is given by
\end{comment}
\begin{align}\label{eq:spectralrand}
\EE{\norm{I-\alpha A_t}^2}=1-\alpha \rhos{P}, {\norm{I-\alpha A_P}^2}=1-\alpha \rhod{P}
\end{align}
Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:dist} to the spectral norms matrices $H_t$ and $H_P$. For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$. If $P$ is Hurwitz, we can argue on similar lines by first transforming it into a positive definite problem $P_U$ and replacing $\rhos{P}$ and $\rhod{P}$ by $\rhos{P_U}$ and $\rhod{P_U}$, and introducing $\cond{U}$ to account for the forward ($\gamma=U^{-1}\theta$) and reverse ($\theta=U\gamma$) transformations using $U^{-1}$ and $U$ respectively.
\item \textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
\item \textbf{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms. The first term is the bias term given by $\B=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$.  The \emph{bias} term that captures the rate at which the initial condition $\norm{\theta_0-\ts}^2$ is forgotten. The second term is the \emph{variance} given by $\V=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}} \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $. The variance term that captures the rate at which noise is rejected.
\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term. This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower. However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:rate} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$. In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other. For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
\item \textbf{Lower Bound} in \Cref{th:lb} shows that
\end{itemize}
\input{illustrate}