### 2017-05-11 plts.tex

\FloatBarrier
\begin{figure}[htp]
\resizebox{\columnwidth}{!}{
\begin{tabular}{ccc}
\begin{tikzpicture}[scale=0.3,font=\Large]
    \begin{axis}[
        xlabel=$t$,
         ylabel=$\norm{\thh_t-\ts}^2$, legend style={at={(0.5,-0.1)},anchor=north}
]
    \addplot[only marks,mark=square,red] plot file {./exp/as};
    \end{axis}
    \end{tikzpicture}
&
\begin{tikzpicture}[scale=0.3,font=\Large,]
\begin{axis}[
xlabel=$t$,
ylabel=$\norm{\thh_t-\ts}^2$, legend style={at={(0.5,-0.1)},anchor=north}
]
\addplot[only marks,mark=square,red] plot file {./exp/pd};
\end{axis}
\end{tikzpicture}

&
\begin{tikzpicture}[scale=0.3,font=\Large]
\begin{axis}[
xlabel=$t$,
ylabel=$\norm{\thh_t-\ts}^2$, legend style={at={(0.5,-0.1)},anchor=north}
]
\addplot[only marks,mark=square,red] plot file {./exp/spd};
\end{axis}
\end{tikzpicture}
\end{tabular}
}
\end{figure}