### 2017-05-19 tune.tex

\FloatBarrier
\begin{figure}[htp]
\resizebox{\columnwidth}{!}{
\begin{tabular}{cc}
\begin{tikzpicture}[scale=0.7]
\begin{axis}[
xlabel=$\sigma_A$,
ylabel=$\alpha_P$, legend pos= north east
]
\addplot[only marks, blue, mark=*, mark options={blue}, error bars/.cd,y dir=both,  y explicit] table [x index=0, y index=1, y error index=2]{./exp/hur};
\addplot[only marks, red, mark=square, mark options={red}, error bars/.cd,y dir=both,  y explicit] table [x index=0, y index=1, y error index=2]{./exp/hur_act};
\addlegendentry{\Cref{alg:tuning}}
\addlegendentry{closed form computation}
\end{axis}
\end{tikzpicture}

&
\begin{tikzpicture}[scale=0.7]
\begin{axis}[
xlabel=$t$,
ylabel=$\normsm{\thh_t-\ts}^2$, legend pos=north east
]

\addplot[only marks,mark=square,red,each nth point=25] plot file {./exp/hur1};
\addplot[only marks,mark=diamond,blue,each nth point=25] plot file {./exp/hur2};
\addplot[only marks,mark=star,green,each nth point=25] plot file {./exp/hur3};
\addplot[only marks,mark=o,black,each nth point=25] plot file {./exp/hur4};
\addplot[only marks,mark=triangle,magenta,each nth point=25] plot file {./exp/hur5};
\addlegendentry{$\sigma_A=0$}
\addlegendentry{$\sigma_A=2$}
\addlegendentry{$\sigma_A=5$}
\addlegendentry{$\sigma_A=10$}
\addlegendentry{$\sigma_A=20$}

\end{axis}
\end{tikzpicture}

\end{tabular}
}
\caption{The left plot shows comparison of the constant step-size $\alpha_P$ as function of $\sigma_A$ found by \Cref{alg:tuning} versus the constant step-size computed in closed form.  The right plot shows the performance of LSA with CS-PR (with the step-size choosen by \Cref{alg:tuning}) for various $\sigma_A$ values.}
\end{figure}